### 2023-02-10 article.tex
Streamline

To determine how useful code is, it needs to be available.
However, it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where 37 out of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
For arcticles in Science magazine (years 2011-2012), 24 out of 204
studies published the code \cite{stodden2018empirical}.
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

@@

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case, the concentration of IL6RA).
To be more precise, it was code that read the data,
subsetted the data, removed outliers, 
performed the statistics and generated the plots.
All these steps are described in the paper (\cite{ahsan2017relative}), 
but the code itself is unpublished
and the paper does not mention any way to obtain it.
Sending an email is known to be unlikely to work 
(see section \ref{sec:code-availability}).

@@

The reason behind paradata is, among others,
to assure reproducible research \cite{huvila2022improving}.
Reproducibility (i.e. to reproduce the same results) 
and replicability (i.e. to reconclude a conclusion)
are fundamental characteristics of scientific studies \cite{patil2019visual}.
For computational science, theoretically, it should be relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In practice, however, it appears that 
the academic culture to reproduce results 
has been lost over time in general \cite{peng2011reproducible},
with labs that embrace reproducibility (for example, \cite{barba2016hard})
are the exception.
To counter the lack of published code, 
it has been suggested to make the reproduction of research 
a minimal requirement for publication \cite{peng2011reproducible}.

@@

For purposes of reproducing (same code, same data) or replicating (same code,
different data) a study, the minimal amount of documentation
is determined by software development best practices.
When an independent researcher reproduces or replicates an experiment, 
it is possible (even ideal!) that no code is read at all.
Due to this, one could argue that no documentation is needed.
However, all code should be documented 'adequately' \cite{peng2006reproducible},
ideally writing code in such a way that it becomes 
self-explanatory \cite{wilson2014best}
and for the remaining code to document the reasons behind it, 
its design and its purpose \cite{wilson2014best}.

@@

Code hosting websites such as GitHub help developers to follow
best practices. One of these is to add a software license \cite{jimenez2017four},
so that it is clear that the software can be reused.

@@

@@

Hosted code commonly has a commit history. For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.
It is a general recommendation to put version control
on all human-produced data \cite{wilson2014best},
as well as openly working on the code from the start \cite{jimenez2017four}.

@@

Publishing code may be disadvantageous for an author.
For science, code should be published, 
this allows reproducible research 
(again, see \cite{haibe2020importance} for a tragic example).
For an author, publishing code alongside an experiment opens up
the possibily to receive questions regarding that code.
Note, however, that not publishing code may put 
oneself in the focus of attention
and -after much effect by others reproducing an incorrect result-
at the cost of a scientific career \cite{baggerly2009deriving}.

>>>>

To determine how useful code is, it needs to be available.
However, it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where 37 out of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
For articles in Science magazine (years 2011-2012), 24 out of 204
studies published the code \cite{stodden2018empirical}.
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

@@

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case, the concentration of IL6RA).
To be more precise, it was code that read the data,
subsetted the data, removed outliers, 
performed the statistics and generated the plots.
For the rest of the discussion, we assume that the
code is available to us (note that it is not
published with the paper, see section \ref{sec:code-availability} 
for a glum estimate of the chance of obtaining the code).

@@

The reason behind paradata is, among others,
to assure reproducible research \cite{huvila2022improving}.
Reproducibility (i.e. to reproduce the same results) 
and replicability (i.e. to reconclude a conclusion)
are fundamental characteristics of scientific studies \cite{patil2019visual}.
For computational science, it seems to be relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In practice, however, it appears that 
the academic culture to reproduce results 
has been lost over time in general \cite{peng2011reproducible},
with labs that embrace reproducibility (for example, \cite{barba2016hard})
are the exception.
One suggested way forward is to make the reproduction of 
research a minimal requirement for publication \cite{peng2011reproducible}.

@@

For purposes of reproducing (same code, same data) or replicating (same code,
different data) a study, the minimal amount of documentation
is determined by software development best practices.
When an independent researcher reproduces or replicates an experiment, 
it is possible (even ideal!) that no code is read at all.
However, the use of the code should be documented,
as 57\% out of 56 Science magazine corresponding authors 
that shared their code
has documented their code well enough to run it 
and understand what it ought to do \cite{stodden2018empirical}.
When one is able to run the code, it could argued 
that no further documentation is needed.
However, all code in general should be documented 
'adequately' \cite{peng2006reproducible},
ideally writing code in such a way that it becomes 
self-explanatory \cite{wilson2014best}
and for the remaining code to document the reasons behind it, 
its design and its purpose \cite{wilson2014best}.

@@

Code hosting websites such as GitHub help developers to follow
best practices. One of these is to add a software license \cite{jimenez2017four},
so that it is clear that the software can be reused.
About one in three of 56 Science magazine corresponding authors 
that shared their code via a URL or email
do not supply a software license \cite{stodden2018empirical}.

@@

For those 22 (out of 180) Science magazine corresponding authors 
that supplied code and that code was estimated to be 
reproducible, one in three had the results reproduced with 
minimal or no effort \cite{stodden2018empirical}.

@@

Hosted code commonly has a commit history,
where a commit is the description of the change, 
as well as the changes in one or more file.
For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.
It is a general recommendation to put version control
on all human-produced data \cite{wilson2014best},
as well as openly working on the code from the start \cite{jimenez2017four}.
However, half of 56 Science magazine corresponding authors 
does not have such a version control mechanism \cite{stodden2018empirical}.

@@

Publishing code may be disadvantageous for an author.
For science, code should be published, 
this allows reproducible research 
(see \cite{haibe2020importance} for a tragic example).
For an author, publishing code alongside an experiment opens up
the possibily to receive questions regarding that code.
Note, however, that not publishing code may put 
oneself in the focus of attention
and -after much effect by others reproducing an incorrect result-
at the cost of a scientific career \cite{baggerly2009deriving}.
